
First, download the relevant libraries
```{r libs, message=FALSE}

library(stringr) # for regular expression work later
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
library(DT)
```

Next, download the relevant tables from Neotoma
```{r download-tables}
tables = c("datasets","datasetdatabases","constituentdatabases","datasettypes")

for (q in seq(length(tables))) {
  tablename=tables[q]
  
  length_table = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=true")))$data
  
  divider = 80000
  num_reps = ceiling(as.numeric(length_table[[1]]$count)/divider)
  
  data_list = list()
  for (i in seq(num_reps)) {
    start= (i-1)*divider
    end = divider
    data_list_portion = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=false&offset=",start,"&limit=",end)))$data
    
    if (!is.null(data_list_portion)) {
      data_list = append(data_list,data_list_portion)
    }
  }
  
  
  data_mat = matrix(nrow=length(data_list),ncol=length(data_list[[1]]))
  
  for (i in seq(length(data_list))) {
    for(j in seq(length(data_list[[1]])) ) {
      if (!is.null(data_list[[i]][[j]])) {
        data_mat[i,j] = data_list[[i]][[j]]
      }}}
  
  
  data_df = as.data.frame(data_mat)
  
  
  names(data_df) =names(data_list_portion[[1]])
  
  assign(paste0(tablename,"_df"), data_df)}
```

Next, I filter for just the vertebrate fauna datasets. Then I break all the distinct datasetids into four big strings, and I use a Neotoma API to get all the data associated with those datasets. And then I do some looping to turn the output lists into dataframes.

``` {r make-verts-data}
vert_datasets = datasets_df %>% left_join(datasettypes_df, by=join_by(datasettypeid)) %>%
  dplyr::filter(datasettype=="vertebrate fauna")

datasets_string1 = paste0(vert_datasets$datasetid[1:1340], collapse=",")
datasets_string2 = paste0(vert_datasets$datasetid[1341:2680], collapse=",")
datasets_string3 = paste0(vert_datasets$datasetid[2681:4020], collapse=",")
datasets_string4 = paste0(vert_datasets$datasetid[4021:5337], collapse=",")


dat1 = content(GET(paste0("https://api.neotomadb.org/v2.0/data/downloads/",datasets_string1,"?limit=9999")))$data
dat2 = content(GET(paste0("https://api.neotomadb.org/v2.0/data/downloads/",datasets_string2,"?limit=9999")))$data
dat3 = content(GET(paste0("https://api.neotomadb.org/v2.0/data/downloads/",datasets_string3,"?limit=9999")))$data
dat4 = content(GET(paste0("https://api.neotomadb.org/v2.0/data/downloads/",datasets_string4,"?limit=9999")))$data


all_dat = dat1 %>% append(dat2) %>% append(dat3) %>% append(dat4)


## This first first loop is just to get the number of rows I'll need
## for my vert_data_mat table
counter=0
for (i in seq(length(all_dat))) {
  for (j in seq(length(all_dat[[i]]$site$collectionunit$dataset$samples))) {
    for (k in seq(length(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum))) {
      counter = counter + 1
    }
  }
}

vert_data_mat = matrix(nrow=counter, ncol=17)


counter2 = 0
for (i in seq(length(all_dat))) {
  for (j in seq(length(all_dat[[i]]$site$collectionunit$dataset$samples))) {
    for (k in seq(length(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum))) {
      counter2 = counter2 + 1
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$datasetid)) {
        vert_data_mat[[counter2,1]] = all_dat[[i]]$site$collectionunit$dataset$datasetid
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$depth)) {
        vert_data_mat[[counter2,2]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$depth
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$sampleid)) {
        vert_data_mat[[counter2,3]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$sampleid
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$thickness)) {
        vert_data_mat[[counter2,4]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$thickness
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$samplename)) {
        vert_data_mat[[counter2,5]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$samplename
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$igsn)) {
        vert_data_mat[[counter2,6]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$igsn
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$analysisunitid)) {
        vert_data_mat[[counter2,7]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$analysisunitid
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$units)) {
        vert_data_mat[[counter2,8]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$units
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$value)) {
        vert_data_mat[[counter2,9]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$value
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$context)) {
        vert_data_mat[[counter2,10]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$context
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$element)) {
        vert_data_mat[[counter2,11]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$element
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$taxonid)) {
        vert_data_mat[[counter2,12]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$taxonid
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$symmetry)) {
        vert_data_mat[[counter2,13]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$symmetry
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$taxongroup)) {
        vert_data_mat[[counter2,14]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$taxongroup
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$elementtype)) {
        vert_data_mat[[counter2,15]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$elementtype
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$variablename)) {
        vert_data_mat[[counter2,16]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$variablename
      }
      
      if(!is.null(all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$ecologicalgroup)) {
        vert_data_mat[[counter2,17]] = all_dat[[i]]$site$collectionunit$dataset$samples[[j]]$datum[[k]]$ecologicalgroup
      }
      
    }
  }
}

all_dat_df = as.data.frame(vert_data_mat)


names(all_dat_df) = c("datasetid","depth","sampleid","thickness","samplename","igsn","analysisunitid","units","value","context","element","taxonid","symmetry","taxongroup","elementtype","variablename","ecologicalgroup")


all_dat_df = all_dat_df %>% left_join(datasetdatabases_df,by=join_by(datasetid)) %>% left_join(constituentdatabases_df,by=join_by(databaseid))
```

Now I clean up the taxa names in various ways, and I try to retain information on the clean ups I'm doing. 

The first warning below is important: in 7 rows, there are more than one slash involved, so I'm losing some taxonomic information there.
```{r clean-data}

## var_cfgenus_remove is the version of variable name with cf. and ? removed from the beginnings
all_dat_df$var_cfgenus_remove <- gsub("^(\\?|cf\\.?|\\?\\s*cf\\.?)\\s*", "", all_dat_df$variablename)

##cf_or_q is a field that says whether the original version of a name had a cf. or question mark at the beginning
all_dat_df = all_dat_df %>% dplyr::mutate(cf_or_q = case_when(var_cfgenus_remove != variablename ~ 1,TRUE~ 0))


## var_spp_remove removes sp. and spp. from the end of the variablename
all_dat_df$var_spp_remove <- gsub("\\s+(sp\\.|spp\\.)$", "", all_dat_df$var_cfgenus_remove) 


##the sp_spp field says whether a variablename originally had sp. or spp. at the end
all_dat_df = all_dat_df %>% dplyr::mutate(sp_spp = case_when(
  var_cfgenus_remove != var_spp_remove ~ 1,TRUE~ 0))

## all_dat_df_split splits up any taxonomic IDs joined by a slash (/) into separate observations. For the second variable, it's sometimes written out like Canis lupus/C. dirus, so I also make sure to replace the second C. with the original genus name
all_dat_df_split = separate(all_dat_df,var_spp_remove,into=c("var1","var2"),sep="/") %>%
  mutate(
    var2 = if_else(
      str_detect(var2, "^[A-Z]\\."),                               # if taxon_clean starts with "X."
      str_replace(var2, "^[A-Z]\\.", str_extract(var1, "^[^ ]+")), # replace "X." with first word of var1
      var2
    )
  )


split_begin = all_dat_df_split %>% dplyr::select(var1,cf_or_q,sp_spp,databasename)
names(split_begin)[1] = "taxon_clean"
split_end = all_dat_df_split %>% dplyr::select(var2,cf_or_q,sp_spp,databasename)
names(split_end)[1] = "taxon_clean"

all_dat_df_split = split_begin %>% rbind(split_end) %>% drop_na(taxon_clean)


## This last cleaning is to find whether there are cfs. after the genus name but before the species name and remove those
all_dat_df_split  = all_dat_df_split  %>%
  mutate(
    cf_species = str_detect(taxon_clean, "cf\\. [A-Z]\\."),                          # TRUE if cf. X. is present
    taxon_clean = str_replace(taxon_clean, "cf\\. [A-Z]\\.[ ]?", "")                      # remove cf. X. and the following space
  )
```

Lastly, I load up the big MDD table and I start to match to it, first at the level of species, then genus, then family, then order. I combine all the pieces and use that at the end to make some visualizations.

```{r match}
mdd = read.csv("Species_Syn_v2.2.csv") %>% distinct(MDD_original_combination,.keep_all=TRUE)


## harmonization by MDD species
harm_by_species = all_dat_df_split %>% left_join(mdd,by=join_by("taxon_clean"=="MDD_original_combination")) %>% drop_na(MDD_order) %>% select(taxon_clean,MDD_genus,MDD_order,MDD_family,cf_or_q,sp_spp,cf_species,databasename) %>% dplyr::mutate(certainty="species")

harm_by_species_na = all_dat_df_split %>% left_join(mdd,by=join_by("taxon_clean"=="MDD_original_combination")) %>% dplyr::filter(is.na(MDD_order))

harm_by_species_distinct_na = harm_by_species_na %>% distinct(taxon_clean)


## harmonization by MDD genera
mdd_genera = mdd %>% distinct(MDD_genus,.keep_all=TRUE)

harm_by_genus = all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_species_distinct_na$taxon_clean) %>% left_join(mdd_genera,by=join_by("taxon_clean"=="MDD_genus")) %>% drop_na(MDD_order)  %>% dplyr::mutate(MDD_genus = taxon_clean) %>% select(taxon_clean,MDD_genus,MDD_order,MDD_family,cf_or_q,sp_spp,cf_species,databasename) %>% dplyr::mutate(certainty="genus")


harm_by_genus_na = all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_species_distinct_na$taxon_clean) %>% left_join(mdd_genera,by=join_by("taxon_clean"=="MDD_genus")) %>% dplyr::filter(is.na(MDD_order)) 

harm_by_genus_distinct_na = harm_by_genus_na %>% distinct(taxon_clean)

## harmonization by MDD family
mdd_families = mdd %>% distinct(MDD_family,.keep_all=TRUE)

harm_by_family = all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_genus_distinct_na$taxon_clean) %>% left_join(mdd_families,by=join_by("taxon_clean"=="MDD_family")) %>% drop_na(MDD_order)  %>% dplyr::mutate(MDD_family = taxon_clean) %>% select(taxon_clean,MDD_genus,MDD_order,MDD_family,cf_or_q,sp_spp,cf_species,databasename) %>% dplyr::mutate(certainty="family")


harm_by_family_na =  all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_genus_distinct_na$taxon_clean) %>% left_join(mdd_families,by=join_by("taxon_clean"=="MDD_family")) %>% dplyr::filter(is.na(MDD_order))

harm_by_family_distinct_na = harm_by_family_na %>% distinct(taxon_clean)


## harmonization by MDD orders
mdd_orders = mdd %>% distinct(MDD_order,.keep_all=TRUE)

harm_by_order = all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_family_distinct_na$taxon_clean) %>% left_join(mdd_orders,by=join_by("taxon_clean"=="MDD_order")) %>% drop_na(MDD_species)  %>%
  dplyr::mutate(MDD_order = taxon_clean) %>% select(taxon_clean,MDD_genus,MDD_order,MDD_family,cf_or_q,sp_spp,cf_species,databasename) %>% dplyr::mutate(certainty="order")


harm_by_order_na =  all_dat_df_split %>% dplyr::filter(taxon_clean %in% harm_by_family_distinct_na$taxon_clean) %>% left_join(mdd_orders,by=join_by("taxon_clean"=="MDD_order")) %>% dplyr::filter(is.na(MDD_species))

harm_by_order_distinct_na = harm_by_order_na %>% distinct(taxon_clean)


all_harm = harm_by_species %>% rbind(harm_by_genus) %>% rbind(harm_by_family) %>% rbind(harm_by_order)


##1790 unmatched taxa in harm_by_species
##1634 in harm_by_order older
#1396 in harm_by_order newer


all_harm$MDD_order <- factor(all_harm$MDD_order, levels = c("Rodentia", "Carnivora", "Artiodactyla","Lagomorpha","Eulipotyphla","Chiroptera","Perissodactyla","Didelphimorphia","Proboscidea","Cingulata","Pilosa","Sirenia","Primates"))

ggplot(all_harm) +
  geom_bar(mapping=aes(x=MDD_order,fill=databasename)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

famharm = all_harm %>% group_by(MDD_order) %>% dplyr::filter(certainty != "order") %>% distinct(MDD_family) %>% count() %>% arrange(desc(n))

names(famharm) = c("order","number families")

genharm = all_harm %>% group_by(MDD_order) %>% dplyr::filter(!certainty %in% c("order","family")) %>% distinct(MDD_genus) %>% count() %>% arrange(desc(n))

names(genharm) = c("order","number genera")


specharm = all_harm %>% group_by(MDD_order) %>% dplyr::filter(!certainty %in% c("order","family","genus")) %>% distinct(taxon_clean) %>% count() %>% arrange(desc(n))

names(specharm) = c("order","number species")

famharm %>% left_join(genharm) %>% left_join(specharm) %>% datatable(rownames=FALSE,options = list(pageLength = 12))
```

